---
title: "Reproducible Research Week 2 Assignment"
author: "Michael Arthur"
date: "July 22, 2018"
output: github_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```


# Part 1 
## Loading and preprocessing the data

1. Load the data
```{r Data Load}

###
activityData <- read.csv("./data/activity.csv")
str(activityData)
```

# Part 2
## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r}
stepsPerDay <- 
      activityData %>% 
      na.omit() %>% 
      group_by(date) %>% 
      summarise(steps = (sum(steps)))
```


2. Histogram of the total number of steps
```{r Histogram}
##histogram
hist(stepsPerDay$steps,
      main = "Frequency of Steps Per Day",
      xlab = "Total Steps")
```

3. Calculate the mean and median of the total steps taken per day

```{r}
##mean
mean(stepsPerDay$steps)
##median
median(stepsPerDay$steps)
##total
sum(stepsPerDay$steps)
```

# Part 3
## What is the average daily activity pattern
1. Make a time series plot of the 5-minute interval and the average number of steps taken, averaged across all days

```{r}
activityPattern <- activityData %>% 
  group_by(interval) %>% 
  na.omit() %>% 
  dplyr::summarize(average = mean(steps))
plot(activityPattern, type ="l")
```

2. Which 5-minute interval, on average across all the days contains the maximum number of steps?

```{r}
activityPattern %>% 
  arrange(desc(average)) %>% 
  top_n(1)
```

#Part 4
## Imputing missing values

1. Calculate and report the number of missing values in the dataset
```{r}
##Imput missing values
sum(is.na(activityData))
```

2. Devise a strategy for filling in all the missing values in the dataset:
* Replace NA's with the average for the interval across all days

3. Create a new dataset that is equal to the original dataset but with the missing data filled in
```{r}
### Replace NA's with average for the interval
activityData.f <- activityData %>% 
  group_by(interval) %>% 
 mutate(steps = ifelse(test = is.na(steps), yes = mean(x = steps, na.rm = TRUE), no = steps))
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
stepsPerDay.f <- 
  activityData.f %>% 
  group_by(date) %>% 
  summarise(steps = (sum(steps)))

hist(stepsPerDay.f$steps,
     main = "Frequency of Steps Per Day (filled)",
     xlab = "Total Steps")

##mean
mean(stepsPerDay.f$steps)
##median
median(stepsPerDay.f$steps)
##total
sum(stepsPerDay.f$steps)
```
Do these values differ from the estimates from the first part of the assignment? 
* The mean does not, the median does slightly slightly here are the differences:
***
Mean
```{r}
##mean
mean(stepsPerDay$steps) - mean(stepsPerDay.f$steps)
```
Median
```{r}
##median
median(stepsPerDay$steps) -
  median(stepsPerDay.f$steps)
```

What is the impact of imputing missing data on the estimates of the total daily number of steps?
* It shifts the median slightly to the left

# Part 5
## Are there differences between weekends and weekdays

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
activityWeekend <- activityData.f %>% 
  mutate(weekday = weekdays(ymd(date))) %>% 
  mutate(DayType = as.factor(ifelse(weekday %in% c("Saturday", "Sunday"), "Weekend", "Weekday")))
```

2. Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

```{r}
a  <- ggplot(data = activityWeekend, 
       aes(x = interval, y= steps)) +
      geom_line()


a + facet_grid(DayType ~ .) +
  theme_minimal()
```
